---
title: "Creating graphs with ggplot2"
author: "Joni"
date: "1/27/2020"
output: html_document
---

# Plots with ggplot2

We are going to make plots in R using the ggplot2 package fromt he tidyverse.

We will install these packages:

- tidyverse: suite of packages for wrangling &viz
- readxl: to read in excel files 
- here: to help with filepaths

```{r setup, message=FALSE, warning=FALSE}
library (tidyverse) #install.packages ("tidyverse")
library (readxl) #install.packages ("readxl")
library(here) # install.packages ("here")
```

#California Naitona Parks Data
```{r read-data, message=FALSE}

ca_np <- read_csv (here("data/ca_np.csv"))
# also be written:
#ca_np <- read_csv ("data/ca_np.csv")

ci_np <- read_excel(here("data", "ci_np.xlsx"))
```
### Make my own first graph!

We're going to make a line plot of year (x-vairable)vs visitors (y-vairable) in ggplot:
Alt+I
```{r}

ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_line()

```

# Update the geom_ to geom_point(), geom_col(), geom_area()


Make a base layer for my CI plots:
```{r}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))

gg_base + geom_point()

```

### Now let's start customizing!

_ Line / point colors: color
- Change line type: linetype
- Change point style: pch
- Change transparency: alpha

```{r}
gg_base + 
  geom_line(
    color = "purple",
    linetype = "dashed"
  )


```

Activity: create a ggplot scatterplot (geom_point) that has triangel symbols (hint: ?pch), of a non-default size (size = #), and that are an updated color

```{r}
gg_base +
  geom_point( 
    color = "dodgerblue",
    pch = 2,
    size = 4
    )

```

### Update graph aesthetics based on vairables

```{r}
gg_base +
  geom_point (
    aes(size = visitors,
        color = year), 
    pch = 10
  )
# Add an argument to update the color based on the year value  
```

### Use gg-themes for overall updates

- theme_light()
- theme_minimal()
_ theme_bw()


```{r}
gg_base +
  geom_area() +
  theme_classic()
```


### Create a multi-seires line plot using data for all CA Naitonal Parks

```{r}
ggplot (data = ca_np, aes(x = year, y = visitors, group = park_name)) +
  geom_line(aes (color = park_name))
# Try to update the line colors based on the park_name variable
```

```{r}
my_np_graph <- ggplot(data = ca_np, aes(x = year, y = visitors)) +
  geom_point(aes(color = park_name), show.legend = FALSE) +
  facet_wrap(~park_name, scales = "free")+
  theme_bw()#~ means as a function of, free means each graph has its own scale
```
Export my graph as a .png
```{r}
ggsave(plot = my_np_graph, here("figures", "np_graph.png"))
```

