---
title: "Tidying session"
author: "Joni"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

# Attaching packages
library(tidyverse)
library(janitor)
library(here)
library(readxl)
```

### Read in inverts.xlsx

```{r}
inverts <- read_excel(here("data","inverts.xlsx"))
```

### `pivot_longer()`
Use single colon (:) to indicate aseries of sequential columns. 

```{r}
inverts_long <- pivot_longer(data = inverts, 
                             cols = '2016' : '2018',
                             names_to = "year",
                             values_to = "sp_count")

#What if you want some in series and some Not in series?
# c('2016':'2018', '2020')
```

What if we want to get it back into wide format?
```{r}
inverts_wide <- inverts_long %>%
  pivot_wider(names_from = common_name, 
              values_from = sp_count)
```
### `janitor:: clean_names()`


```{r}
inverts_wide <- inverts_wide %>%
  clean_names()
```
### `tidyr:: unite()` to combine information from columns

Examples: merge the site code and year into a single column called "site_year"
```{r}
inverts_unite <- inverts_long %>%
  unite(col = "site_year",
        c(site, year),
        sep = "_")

```
### kActivity:

Create a new object called 'inverts_moyr' staring from inverts_long that unites the month and year columns into a single column named "mo_yr", using a slash "/" as the separator.. then try some other fun separator

```{r}
inverts_moyr <- inverts_long %>%
  unite(col = "mo_yr",
        c(month, year),
        sep = "/")

```
### `tidyr:: separate()`
```{r}
inverts_sep <- inverts_unite %>%
  separate(col = site_year,
           into = c("my_site", "my_year"))
```

### `stringr` +`mutate()`

mutate: add a column or update a column. better to add a column than update

Example: using built-in 'rees' dataset with mutate() to add 5 to all of the tree heights

```{r}
trees_new <- trees %>% 
  mutate(height_update = Height + 5)
```
### `str_replace()`
replace places in California under common_name with CA
```{r}
ca_abbr <- inverts %>%
  mutate(common_name = 
          str_replace(common_name, 
                      pattern = "california",
                      replacement = "CA") )
```

Example: In the common_name column, you want to replace the word "spiny" with the word "awesome"

store as `awesom_lobsters`

```{r}
awesome_lobsters <- inverts %>%
  mutate(common_name = 
          str_replace(common_name, 
                      pattern = "spiny",
                      replacement = "awesome") )
```

